Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}

### 2023-02-21 draft/draft.tex
lots of updates!

\title{Pollution, agricultural productivity, and development: Evidence from coal plants in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-02-11}

@@

\caption{\label{tab:plantresultstable}Local characteristics and the construction of coal plants}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1991 census} & \multicolumn{2}{c}{2001 census} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & 1991 & 2001 & 2001 & 2011\\
\midrule
pop (log) & 0.068*** & 0.0003 & 0.009*** & 0.002***\\
 & (0.0002) & (0.0002) & (0.0003) & (0.0002)\\
literacy (prop) & 0.127*** & -0.005*** & 0.043*** & 0.030***\\
 & (0.002) & (0.001) & (0.002) & (0.002)\\
has paved road & -0.036*** & 0.003*** & -0.008*** & -2.68e-7\\
 & (0.0008) & (0.0004) & (0.0008) & (0.0006)\\
ag productivity &  &  & 0.033*** & 0.005***\\
 &  &  & (0.0005) & (0.0004)\\
sub-sample & all & no plant & all & no plant\\
\midrule
observations & 490,882 & 447,288 & 516,798 & 476,517\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & (1) & (2) & (3) & (4)\\
\midrule
wind & 0.045*** & 0.014*** & 0.063*** & 0.015***\\
 & (0.004) & (0.001) & (0.005) & (0.002)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village & Yes & Yes & Yes & Yes\\
month & Yes & No & Yes & No\\
district-month & No & Yes & No & Yes\\
\midrule
observations & 22,345,092 & 22,345,092 & 14,896,728 & 14,896,728\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

In addition to serving as an example, Table \ref{tab:pollutiontable} also helps validate the use of wind direction as a proxy for exposure to pollution. The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated. Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher. The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.01 and 0.06 \(\mathrm{\mu g/m^3}\).

@@

\caption{\label{tab:yieldtable}Wind direction and agricultural productivity}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
  & (1) & (2) & (3) & (4) & (5)\\
\midrule
wind & -0.003*** & -0.003*** & -0.0007*** & -0.002*** & -0.004***\\
 & (0.0002) & (0.0002) & (9.24e-5) & (0.0002) & (0.0003)\\
rain (z) &  & 0.029*** & 0.008*** & 0.075*** & 0.006***\\
 &  & (0.0004) & (0.001) & (0.002) & (0.0005)\\
mean temp (10s) &  & -0.0004*** & -0.007*** & -0.002*** & -0.008***\\
 &  & (5.53e-5) & (0.0004) & (0.0001) & (9.83e-5)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & No & No\\
year & Yes & Yes & No & Yes & Yes\\
district-year-season & No & No & Yes & No & No\\
village & No & No & No & Yes & Yes\\
\midrule
observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
  & (1) & (2) & (3) & (4) & (5)\\
\midrule
particulate matter & 0.0000 & 0.0005*** & -0.0004*** & 0.002*** & 0.0004***\\
(PM 2.5, '000s) & (0.0000) & (3.17e-5) & (5.72e-5) & (9.87e-5) & (3.7e-5)\\
rain (z) &  & 0.030*** & 0.008*** & 0.075*** & 0.007***\\
 &  & (0.0004) & (0.001) & (0.002) & (0.0005)\\
mean temp (10s) &  & -0.0003*** & -0.007*** & -0.002*** & -0.008***\\
 &  & (5.63e-5) & (0.0004) & (0.0001) & (9.79e-5)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & No & No\\
year & Yes & Yes & No & Yes & Yes\\
district-year-season & No & No & Yes & No & No\\
village & No & No & No & Yes & Yes\\
\midrule
observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

Of course, it is not wind, per se, that we are interested in. Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity. Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{fig:yieldtable}, with a key difference: instead of wind exposure, it uses average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al. (2020) -- during the season as the key explanatory variable. I rescale particulate matter by dividing by one thousand to help with presentation of the coefficients. Several interesting patterns emerge. In the first column we see no relationship between particulate matter and agricultural productivity, while there is actually a positive relationship in columns two, four, and five. There is a small, negative relationship between pollution and productivity in column three.

@@

The second-stage results first indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity. The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for \(\mathrm{PM_{2.5}}\) point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above. This makes sense, of course, since the IV is essentially scaling up the effects of wind. The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants. The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so. In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since coal plants are placed in more productive areas. The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.

@@

\caption{\label{tab:yieldtabletwo}Pollution and agricultural productivity, IV estimates}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
  & (1) & (2) & (3) & (4) & (5)\\
\midrule
particulate matter & -0.021*** & -0.024*** & -0.032*** & -0.012*** & -0.035***\\
(PM 2.5) & (0.001) & (0.002) & (0.005) & (0.001) & (0.003)\\
rain (z) &  & -0.011*** & -4.24e-5 & 0.078*** & -0.042***\\
 &  & (0.003) & (0.002) & (0.002) & (0.004)\\
mean temp (10s) &  & -0.005*** & -0.007*** & -0.002*** & -0.008***\\
 &  & (0.0003) & (0.0004) & (0.0001) & (0.0001)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & No & No\\
year & Yes & Yes & No & Yes & Yes\\
district-year-season & No & No & Yes & No & No\\
village & No & No & No & Yes & Yes\\
\midrule
observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
\midrule
\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
wind & 0.145*** & 0.111*** & 0.022*** & 0.157*** & 0.105***\\
 & (0.002) & (0.002) & (0.002) & (0.002) & (0.004)\\
rain (z) &  & -1.68*** & -0.239*** & 0.225*** & -1.37***\\
 &  & (0.011) & (0.017) & (0.015) & (0.010)\\
mean temp (10s) &  & -0.192*** & 0.002 & -0.021*** & -0.007***\\
 &  & (0.0009) & (0.004) & (0.001) & (0.002)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects. Table \ref{tab:yieldtablehet} presents four different analyses: effects of exposure by maximum exposure, multiplicative effects of shocks, the possibility of non-linearities, and the differential effects of exposure based on initial agricultural productivity.

The first two columns split the sample based on maximum exposure; villages with maximum exposure above the median are included in column (1) while those below the median are in column (2). Interestingly, the largest effects are seen on villages in the lower half of the distribution. Apparently, smaller maximum effects lead to larger overall effects of pollution exposure on agricultural productivity. In other words, it is not the places where wind is most likely to blow from a coal plant that experience the largest decrease in productivity in response to an additional day of wind-driven pollution.

@@

\caption{\label{tab:yieldtablehet}Heterogeneity in the effects of pollution on productivity}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{exposure} & \multicolumn{2}{c}{yield} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & >p(50) & <=p(50) & >p(50) & <=p(50) & \\
\midrule
pm & -0.074*** & -0.012*** & -0.006*** & -0.066*** & -0.021***\\
 & (0.004) & (0.002) & (0.001) & (0.006) & (0.001)\\
pm (10s) times rain &  &  &  &  & 0.005***\\
 &  &  &  &  & (0.000)\\
weather & yes & yes & yes & yes & yes\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & Yes & Yes\\
year & Yes & Yes & Yes & Yes & Yes\\
\midrule
observations & 650,770 & 1,720,451 & 1,192,644 & 1,178,577 & 2,371,221\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. The first two columns split the sample based on the *maximum* observed exposure. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

Column three looks add the possible multiplicative effects of rainfall shocks and pollution shocks on agricultural productivity. In previous tables, rainfall always entered the regression positively -- as we would expect -- indicating that higher rainfall in a given season leads to higher agricultural productivity. However, an interesting question is whether rainfall and pollution shocks might interact. The positive coefficient on the interaction term between days of wind and rainfall during the agricultural season indicates that there is a multiplicative effect of shocks; when wind days are higher (bad) and rainfall is lower (bad), the overall negative effects are even worse than when just one shock happens. Considering that climate change is leading to more variable rainfall and more negative rainfall shocks, these results are potentially worrisome, especially for smallholders.

Column four adds a quadratic in wind to see if the negative effects of pollution on agricultural productivity are increase or decreasing in wind. In column four, the coefficient on the linear term is 0.0006 while the coefficient on the quadratic term is -0.0002, meaning that the effect of wind is decreasing for any days of wind over three The negative coefficient also indicates that this negative effect is consistently decreasing, meaning the effect gets worse as average wind days gets higher. Note that this is somewhat different from the regressions in the first two columns, which split the sample by the \emph{maximum} days of wind observed in the sample, not the mean.

In the final column I add to the regression an interaction between wind and initial agricultural productivity (defined in 2002, the first year observed in the productivity data). The linear term drops out of the regression since it is invariant within villages, leaving just the interaction term. The negative coefficient on the interaction term of -0.001 indicates that the effect of pollution is more negative in areas where agricultural productivity is higher. Note that agricultural productivity is defined in logs, so this can be interpreted as a percentage effect, not a level effect. Apparently, wind-driven pollution is particularly detrimental to agricultural productivity in the most productive places.

@@

Although the results with district-year-season help mitigate this concern, Table \ref{tab:yieldtablepostplant} in the appendix checks this possibility in a different way. Specifically, I restrict the sample to only village-years \emph{after} the introduction of a coal plant. This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening. The coefficients in Table \ref{tab:yieldtablepostplant} show that the overall effects are essentially unchanged, with just a small change in coefficient magnitude (driven mostly by rounding).

>>>>

\title{Pollution, agricultural productivity, and development: Evidence from coal plants in India\footnote{I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-02-20}

@@

\caption{\label{tab:plantresultstable}Local characteristics and the construction of coal plants}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1991 census} & \multicolumn{2}{c}{2001 census} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & 1991 & 2001 & 2001 & 2011\\
\midrule
pop (log) & 0.068*** & 0.000 & 0.009*** & 0.002***\\
 & (0.000) & (0.000) & (0.000) & (0.000)\\
literacy (prop) & 0.127*** & -0.005*** & 0.044*** & 0.030***\\
 & (0.002) & (0.001) & (0.002) & (0.002)\\
has paved road & -0.037*** & 0.003*** & -0.008*** & -0.000\\
 & (0.001) & (0.000) & (0.001) & (0.001)\\
ag productivity &  &  & 0.033*** & 0.005***\\
 &  &  & (0.001) & (0.000)\\
sub-sample & all & no plant & all & no plant\\
\midrule
observations & 490,882 & 447,288 & 516,798 & 476,517\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\caption{\label{tab:pollutiontable}Wind direction and particulate matter}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{1998-2015} & \multicolumn{2}{c}{2002-2013} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & (1) & (2) & (3) & (4)\\
\midrule
wind (10s) & 0.002*** & 0.002*** & 0.006*** & 0.001***\\
 & (0.000) & (0.000) & (0.001) & (0.000)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village & Yes & Yes & Yes & Yes\\
month & Yes & No & Yes & No\\
district-month & No & Yes & No & Yes\\
\midrule
observations & 24,376,458 & 24,376,458 & 16,250,973 & 16,250,973\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

In addition to serving as an example, Table \ref{tab:pollutiontable} also helps validate the use of wind direction as a proxy for exposure to pollution. The first two columns include all available years of data while the last two columns restrict estimation only to the years for which the agricultural productivity regressions are estimated. Across all four columns, the story is the same: when the wind is blowing in the direction of a village from a coal plant within 30 kilometers, estimated \(\mathrm{PM_{2.5}}\) is substantially higher. The regressions indicate that one additional day of wind in the direction of the village increases mean \(\mathrm{PM_{2.5}}\) for the month by between 0.00 and 0.00 \(\mathrm{\mu g/m^3}\).

@@

\caption{\label{tab:yieldtable}Wind direction and agricultural productivity}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
  & (1) & (2) & (3) & (4) & (5)\\
\midrule
wind & -0.003*** & -0.003*** & -0.001*** & -0.002*** & -0.004***\\
 & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
rain (z) &  & 0.029*** & 0.008*** & 0.075*** & 0.006***\\
 &  & (0.000) & (0.001) & (0.002) & (0.000)\\
mean temp (10s) &  & -0.000*** & -0.007*** & -0.002*** & -0.008***\\
 &  & (0.000) & (0.000) & (0.000) & (0.000)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & No & No\\
year & Yes & Yes & No & Yes & Yes\\
district-year-season & No & No & Yes & No & No\\
village & No & No & No & Yes & Yes\\
\midrule
observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\caption{\label{tab:yieldtablepm}Particulate matter (PM 2.5) and agricultural productivity}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
  & (1) & (2) & (3) & (4) & (5)\\
\midrule
particulate matter & 0.0000 & 0.0329*** & -0.0787*** & -0.0509*** & 0.0216***\\
(log PM 2.5) & (0.0000) & (0.0028) & (0.0040) & (0.0033) & (0.0052)\\
rain (z) &  & 0.0290*** & 0.0079*** & 0.0760*** & 0.0064***\\
 &  & (0.0004) & (0.0013) & (0.0017) & (0.0005)\\
mean temp (10s) &  & -0.0003*** & -0.0071*** & -0.0020*** & -0.0080***\\
 &  & (0.0001) & (0.0004) & (0.0001) & (0.0001)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & No & No\\
year & Yes & Yes & No & Yes & Yes\\
district-year-season & No & No & Yes & No & No\\
village & No & No & No & Yes & Yes\\
\midrule
observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

Of course, it is not wind, per se, that we are interested in. Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity. Table \ref{tab:yieldtablepm} presents the same regressions as Table \ref{fig:yieldtable}, with a key difference: instead of wind exposure, it uses average concentration of \(\mathrm{PM_{2.5}}\) -- from Hammer et al. (2020) -- during the season as the key explanatory variable.

Several interesting patterns emerge. We only observe a negative relationship between particulate matter and agricultural productivity in column three; the coefficient in column two is actually positive. Breaking out the effects into the monsoon and winter seasons, we also see a positive relationship between pollution and agricultural productivity.

@@

The second-stage results first indicate that \(\mathrm{PM_{2.5}}\) is a very strong predictor of agricultural productivity. The smallest coefficient -- that for the monsoon season in column (4) -- and the mean absolute deviation for \(\mathrm{PM_{2.5}}\) point to a change in agricultural productivity of around 23 percent, which is much higher than the wind results documented above. This makes sense, of course, since the IV is essentially scaling up the effects of wind.

The comparison of the results to those in the previous table also confirm the concerns around endogenous placement of coal plants. The coefficients in Table \ref{tab:yieldtabletwo} are always more negative than those in Table \ref{tab:yieldtablepm}, and sometimes substantially so. In other words, the simple regression of agricultural productivity on pollution results in coefficients on pollution that are upwards biased, since higher levels of pollution occur in more productive areas, on average. The IV strategy corrects this bias and leads to a robust, negative relationship between pollution and agricultural productivity.

@@

\caption{\label{tab:yieldtabletwo}Pollution and agricultural productivity, IV estimates}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{all} & \multicolumn{1}{c}{monsoon} & \multicolumn{1}{c}{winter} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
  & (1) & (2) & (3) & (4) & (5)\\
\midrule
particulate matter & -1.190*** & -1.027*** & -1.671*** & -0.525*** & -1.741***\\
(log PM 2.5) & (0.070) & (0.071) & (0.235) & (0.043) & (0.139)\\
rain (z) &  & 0.028*** & 0.012*** & 0.080*** & 0.001**\\
 &  & (0.000) & (0.001) & (0.002) & (0.001)\\
mean temp (10s) &  & -0.001*** & -0.009*** & -0.002*** & -0.007***\\
 &  & (0.000) & (0.000) & (0.000) & \vphantom{2} (0.000)\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & No & No\\
year & Yes & Yes & No & Yes & Yes\\
district-year-season & No & No & Yes & No & No\\
village & No & No & No & Yes & Yes\\
\midrule
observations & 2,387,417 & 2,371,221 & 2,371,221 & 1,256,951 & 1,114,270\\
\midrule
\textbf{first stage:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
wind & 0.003*** & 0.003*** & 0.000*** & 0.004*** & 0.002***\\
 & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
rain (z) &  & -0.001*** & 0.002*** & 0.009*** & -0.003***\\
 &  & (0.000) & (0.000) & (0.000) & \vphantom{1} (0.000)\\
mean temp (10s) &  & -0.000*** & -0.001*** & 0.001*** & 0.001***\\
 &  & (0.000) & (0.000) & (0.000) & (0.000)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

\hypertarget{heterogeneity}{%
\subsection{Heterogeneity}\label{heterogeneity}}

Having established that exposure to pollution has a negative effect on agricultural productivity, I now move to heterogeneity in these effects. Table \ref{tab:yieldtablehet} presents three different analyses: heterogeneity by median exposure, heterogeneity by initial yield, and possible multiplicative effects of shocks. All regressions use the IV strategy detailed above and I include the F statistic from the first stage regressions (of which there are two in the last column) at the bottom of the table. Table \ref{tab:yieldtablehet1st} in the appendix presents the first-stage results for all columns.

The first two columns split the sample based on median exposure. In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure. The coefficient in the first column is 2.5 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common. One possible explanation is that exposure is cumulative, leading to larger and larger effects. I return to this in column five.

Columns two and three split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003). Column three presents results for villages above the median while column four presents results for villages below the median. The largest effects are apparently concentrated in the lower half of the distribution. Perhaps more productive areas are more likely to have ways to combat the effects of shocks -- e.g.~by adjusting the use of other inputs -- or maybe less productive areas are inherently more susceptible to the effects of pollution, perhaps due to direct effects on land productivity.

Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common. In the last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another. Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.

@@

\caption{\label{tab:yieldtablehet}Heterogeneity in the effects of pollution on productivity}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{1.8cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{exposure} & \multicolumn{2}{c}{yield} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & >p(50) & <=p(50) & >p(50) & <=p(50) & \\
\midrule
PM 2.5 & -1.082*** & -2.523*** & -0.439*** & -1.711*** & -0.809***\\
 & (0.073) & (0.624) & (0.046) & (0.146) & (0.060)\\
PM time rain &  &  &  &  & 0.180***\\
 &  &  &  &  & (0.008)\\
weather & Yes & Yes & Yes & Yes & Yes\\
\textbf{fixed effects:} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
village-season & Yes & Yes & Yes & Yes & Yes\\
year & Yes & Yes & Yes & Yes & Yes\\
\midrule
F (1st stage, PM) & 6,488 & 60 & 2,867 & 2,150 & 2,958\\
F (1st stage, PM times rain) & -- & -- & -- & -- & 1,697\\
observations & 1,051,010 & 1,320,211 & 1,193,764 & 1,177,457 & 2,371,221\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. The first two columns split the sample based on the median exposure (by village). The second two columns split the sample based on initial -- the first year available -- agricultural productivity. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

The coefficient on the level term for pollution is negative. Since rainfall is defined as a z score, this represents the effect of pollution only when rainfall is at its mean value for a given season. The interaction term, on the other hand, is positive. Since higher values of rainfall are generally good, this positive coefficient represents a compounding effect of pollution and rainfall; when rainfall is low (bad), the effects of particulate matter on agricultural productivity is larger (more negative). Another way to think about this, however, is that higher values of rainfall dampen the effects of pollution shocks. This could have very profound effects on welfare, since this essentially increases the variance of agricultural productivity across years (which is already increasing due to climate change).

@@

Although the results with district-year-season help mitigate this concern, Table \ref{tab:yieldtablepostplant} in the appendix checks this possibility in a different way. Specifically, I restrict the sample to only village-years \emph{after} the introduction of a coal plant. This drops any pre-plant-opening observations -- all of which are mechanically zeros for wind -- to prevent endogeneity due to the opening. The coefficients in Table \ref{tab:yieldtablepostplant} show that the overall effects are essentially unchanged across any specifications. Comparing changes in coefficients can also help reduce any effects driven by other polluting industries located near coal plants; since the pre-opening exposure variable is zero, the coefficients in Table \ref{tab:yieldtablepostplant} would likely increase in magnitude if the effects were driven by non-coal pollution.
